---
output: 
  html_document:
    css: 
     - custom.css
     - https://maxcdn.bootstrapcdn.com/bootstrap/3.3.7/css/bootstrap.min.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
set.seed(10000)
library(tidyverse)
library(magrittr)
library(httr)
library(jsonlite)
library(psych)
library(knitr)
```

<div class='jumbotron'>
  <h2 class='display-3 text-uppercase'>Project Proposal</h2>
  <h4 class='right text-uppercase'>By Brian Weinfeld</h4>
  <div class='clearfix'></div>
  <h5 class='right text-uppercase'>March 29, 2018</h5>
</div>

* Data Collection

```{r}
games.data <- GET('https://api.mysportsfeeds.com/v1.2/pull/nhl/2015-2016-regular/full_game_schedule.json', authenticate('njpsy', 'asdfasdf'), 
            add_headers('Content-Type'='application/json', 'Accept-Encoding'='gzip')) %>%
  content(as='text', encoding='UTF-8') %>%
  fromJSON(flatten=TRUE) %>%
  extract2(1) %>%
  extract2(2) %>%
  as.tibble()

game.dates <- games.data %>%
  select(6) %>%
  mutate(date=str_replace(date, '(\\d+)-(\\d+)-(\\d+)', '\\1\\2\\3')) %>%
  unique()

API.Query <- function(date){
  Sys.sleep(1)
  GET('https://api.mysportsfeeds.com/v1.2/pull/nhl/2015-2016-regular/scoreboard.json', authenticate('njpsy', 'asdfasdf'), 
      add_headers('Content-Type'='application/json', 'Accept-Encoding'='gzip'),
      query=list('fordate'=date)) %>%
    content(as='text', encoding='UTF-8') %>%
    fromJSON(flatten=TRUE) %>%
    extract2(1) %>%
    extract2(2) %>%
    as.tibble()
}

box.data <- game.dates$date %>%
  map_df(~API.Query(.)) %>%
  select(8, 24) %>%
  unnest() %>%
  mutate(goals = awayScore %>% as.numeric() + homeScore %>% as.numeric()) %>%
  select(-c(3:6)) %>%
  setNames(c('game.ID'='id', '@number'='period', 'goals'='goals')) %>%
  filter(period %in% 1:3) %>%
  as.tibble()
```

* Research question

In the NHL (National Hockey League), is there a relationship between the period of the game and number of goals scored by either team?

* What are the cases, and how many are there?

Each case represents the number goals scored per period in each game of the 2015-2016 regular hockey season. There are 1230 games from which 100 will be selected via SRS.

* Describe the method of data collection.

Raw data is collected by mysportsfeed.com which maintains an API on all major sports created and supported by fans. I will query the API and collected the required data.

* What type of study is this (observational/experiment)?

This is an observational study.

* Data Source: If you collected the data, state self-collected. If not, provide a citation/link.

The total raw data is collected by [mysportsfeed.com](https://mysportsfeed.com), however I will be querying, cleaning and storing the relevant data.

* Response: What is the response variable, and what type is it (numerical/categorical)?

The response variable is the number of goals scored in the period and is numerical.

* Explanatory: What is the explanatory variable(s), and what type is it (numerical/categorival)?

The explanatory variable is the period of the game and is categorical (1st, 2nd or 3rd)

* Relevant summary statistics

```{r}
box.data %>%
  filter(period == '1') %>%
  select(goals) %>%
  describe()

box.data %>%
  filter(period == '2') %>%
  select(goals) %>%
  describe()

box.data %>%
  filter(period == '3') %>%
  select(goals) %>%
  describe()

ggplot(box.data) +
  geom_bar(aes(x=goals, y=(..count../sum(..count..)), fill=period))  +
  facet_wrap(~period, nrow=3, ncol=1) +
  labs(x='Goals', 
       y='Proportion of Goals Scored',
       fill='Period',
       title='Proportion of Goals Scored')
```